R for Clinical Study Reports and Submission

### 2021-10-27 submission-package.Rmd
use ectddemo + update paths

## The Whole Game

In eCTD deliverable, the analysis dataset and source code are saved 
under the module 5 folder `m5/datasets/<study>/analysis/adam`.

For illustration and preparation purpose, we temporarily save
the under the `vignettes/esub/` folder.

@@

```
vignettes/esub/analysis/adam/
├── datasets
│   ├── adrg.docx
│   ├── analysis-results-metadata.docx
│   ├── define-v1-updated-html.xsl
│   ├── define.pdf
│   └── define.xml
└── programs
    ├── tlf-01-disposition.txt
    ├── tlf-02-population.txt
    ├── tlf-03-baseline.txt
    ├── tlf-04-efficacy.txt
    ├── tlf-05-ae-summary.txt
    ├── tlf-06-ae-spec.txt
    └── r0pkgs.txt
```

@@

Folder path: `vignettes/esub/analysis/adam/datasets/`.

@@

Folder path: `vignettes/esub/analysis/adam/programs/`.

@@

- The filenames under `programs` should all use lower case
(with no underscores or other special characters).
- The `.txt` files should only contain ASCII characters.
This can be verified by `pkglite::verify_ascii()`
- The `.docx` files will be converted to PDF files for formal submission.

Now you have a general idea about the relevant components of the
submission package. 
We will focus on preparing the internal R packages in the following sections.

## Practical Considerations for R Package Submissions

@@

Let's pack the `esubdemo` package into a text file.
Assume the source package path is `~/esubdemo/`,
you should be able to pack the package with a single pipe:

@@

"~/esubdemo/" %>%
  collate(file_ectd(), file_auto("inst")) %>%
  pack(output = "r0pkgs.txt")
```

@@

To pack multiple R packages, simply feed multiple file collection as inputs:

@@

The packages are always packed in the specified order and are always
unpacked and installed in the same order. Therefore, make sure to
pack the low-level dependencies first.

For more details on how to customize file specifications
and operate on file collections,
check out the vignette
[Generate File Specifications](https://merck.github.io/pkglite/articles/filespec.html)
and
[Curate File Collections](https://merck.github.io/pkglite/articles/filecollection.html).

@@

You should always verify if the file only contains ASCII characters:

@@

```{r, eval = FALSE}
pkglite::unpack("r0pkgs.txt", output = "/tmp/", install = TRUE)
```

@@

You can then proceed to save the file `r0pkgs.txt` to the folder
`esub/analysis/adam/programs` (or specify the output text file
path above directly).

## Prepare Analysis Programs for Submission

@@

```{r, eval=FALSE}
knitr::purl(
  "tlf-01-disposition.Rmd",
  output = "esub/analysis/adam/programs/tlf-01-disposition.txt"
)
```

Repeat this for all the R Markdown files.

@@

@@

Specifically, we need to update two sections.
The first is the "Macro Programs" section, where
R and R package versions with a snapshot date are provided.

For example:

@@

Submitted R programs have [specific patterns] in filenames.
All internally developed R functions are saved in the r0pkgs.txt file.
The recommended steps to unpack these R functions for analysis output
programs are described in Appendix.

@@

@@

- Section 2: indicate the **Programming Language**
- Section 3: document the details of the R programs listed in section 3.

>>>>

The demo submission package (not to be confused with the R package above)
is [`ectddemo`](https://github.com/elong0527/ectddemo), also available on GitHub.
You can download or clone it:

```bash
git clone https://github.com/elong0527/ectddemo.git
```

We will assume the paths to the two folders are `esubdemo/` and `ectddemo/` below.

## The whole game

In eCTD deliverable, the analysis dataset and source code are saved 
under the eCTD module 5 (clinical study reports) folder

```
ectddemo/m5/datasets/<study>/analysis/adam/
```

@@

```
ectddemo/m5/datasets/ectddemo/analysis/adam/
├── datasets
│   ├── adae.xpt
│   ├── ...
│   ├── adrg.pdf
│   ├── analysis-results-metadata.pdf
│   ├── define.xml
│   └── define2-0-0.xsl
└── programs
    ├── r0pkgs.txt
    ├── tlf-01-disposition.txt
    ├── tlf-02-population.txt
    ├── tlf-03-baseline.txt
    ├── tlf-04-efficacy.txt
    ├── tlf-05-ae-summary.txt
    └── tlf-06-ae-spec.txt
```

@@

Folder path: `ectddemo/m5/datasets/ectddemo/analysis/adam/datasets/`.

@@

Folder path: `ectddemo/m5/datasets/ectddemo/analysis/adam/programs/`.

@@

- The filenames under `programs/` should all use lower case
(with no underscores or other special characters).
- The `.txt` files should only contain ASCII characters.
This can be verified by `pkglite::verify_ascii()`
- Any `.docx` files should be converted to PDF files for formal submission.

Now you have a general idea about the relevant components of the
submission package. 
We will start preparing the proprietary R packages in the following sections.

## Practical considerations for R package submissions

@@

Let's pack the `esubdemo` package into a text file.
Assume the source package path is `esubdemo/`,
you should be able to pack the package with a single pipe:

@@

"esubdemo/" %>%
  collate(file_ectd(), file_auto("inst")) %>%
  pack(output = "r0pkgs.txt")
```

@@

To pack multiple R packages, simply feed multiple file collections as inputs:

@@

The R packages are always packed in the specified order and are always
unpacked and installed in the same order. Therefore, make sure to
pack the low-level dependencies first.

For more details on how to customize file specifications
and operate on file collections,
check out the vignette
[generate file specifications](https://merck.github.io/pkglite/articles/filespec.html)
and
[curate file collections](https://merck.github.io/pkglite/articles/filecollection.html).

@@

You should always verify if the text file only contains ASCII characters:

@@

```{r, eval = FALSE}
unpack("r0pkgs.txt", output = "/tmp/", install = TRUE)
```

@@

You can then proceed to move the file `r0pkgs.txt` to the folder
`ectddemo/m5/datasets/ectddemo/analysis/adam/programs`,
or specify the output text file path above directly.

## Prepare analysis programs for submission

@@

```{r, eval=FALSE}
input_path <- "esubdemo/vignettes/"
output_path <- "ectddemo/m5/datasets/ectddemo/analysis/adam/programs/"

convert_rmd <- function(filename, input_dir, output_dir) {
  knitr::purl(
    file.path(input_dir, paste0(filename, ".Rmd")),
    output = file.path(output_dir, paste0(filename, ".txt"))
  )
}

"tlf-01-disposition" %>% convert_rmd(input_path, output_path)
"tlf-02-population" %>% convert_rmd(input_path, output_path)
"tlf-03-baseline" %>% convert_rmd(input_path, output_path)
"tlf-04-efficacy" %>% convert_rmd(input_path, output_path)
"tlf-05-ae-summary" %>% convert_rmd(input_path, output_path)
"tlf-06-ae-spec" %>% convert_rmd(input_path, output_path)
```

@@

To automate this process:

```{r, eval=FALSE}
header <- readLines(textConnection("# Note to Reviewer
# To rerun the code below, please refer to the ADRG appendix.
# After the required packages are installed,
# the path variable needs to be defined by using the example code below.
#
# path = list(adam = \"/path/to/esub/analysis/adam/datasets\") # Modify to use actual location
# path$outtable = path$outgraph = \".\" # Outputs saved to the current folder"))

append_header <- function(filename, output_dir, header) {
  file <- file.path(output_dir, paste0(filename, ".txt"))
  x <- readLines(file)
  y <- c(header, "", x)
  writeLines(y, con = file)
  invisible(file)
}

"tlf-01-disposition" %>% append_header(output_path, header)
"tlf-02-population" %>% append_header(output_path, header)
"tlf-03-baseline" %>% append_header(output_path, header)
"tlf-04-efficacy" %>% append_header(output_path, header)
"tlf-05-ae-summary" %>% append_header(output_path, header)
"tlf-06-ae-spec" %>% append_header(output_path, header)
```

@@

Specifically, we need to update two sections in ADRG.

The first section is "Macro Programs", where R and R package versions
with a snapshot date are provided. For example:

@@

Submitted R programs have [specific patterns] in filenames.
All internally developed R functions are saved in the r0pkgs.txt file.
The recommended steps to unpack these R functions for analysis output
programs are described in the Appendix.

@@

An example ADRG following this template can be found in
[ectddemo](https://github.com/elong0527/ectddemo/blob/master/m5/datasets/ectddemo/analysis/adam/datasets/adrg.pdf).

@@

- Section 2: indicate the **Programming Language**;
- Section 3: document the details of the R programs listed in section 3.
